Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-04-10 WRITEUP/ms.tex
minor edits to 6.1

The local sound speed of all identified molecular complexes is much
smaller than their turbulent velocities\footnote{Velocity dispersions
  calculated from velocity field are comparable to those calculated
  from the pressure field 
    \MM{The turbulent pressure field has not been
    mentioned in section 3, so you'll need to add a sentence there
    mentioning its presence and behavior to mention it here}, 
    \DL{AP: also mention turbulent pressure found for the bulk of dense gas in the simulation in the phase-space analysis presented in \citet{Pallottini17a} here?}
indicating that contribution from shear is unlikely to dominate the velocity dispersions of the smaller MCCs identified, especially 
at the highest $n_{\rm ncut}$.\label{ftn:veldisp}}. 
On average, the Mach number of the MCCs identified at the highest density threshold is $\mathcal{M} = \sigma_{\rm NT}/c_s\simeq 50$, 
but this depends on the evolutionary stage of \flower.  % see e.g., test_brute/leaf_fields_ss16-28/18.96_10_fields/Mach_cloud-mass.png
This is consistent with the global analysis done on \flower by \citet{Vallini18a}.
We also compute the Mach number from turbulent and thermal pressure to determine the importance of 
differential rotation in the derived velocity dispersion: 
\begin{equation}
\mathcal{M} \equiv \sqrt{\frac{P_{\rm NT} + P}{ P }}\,,
\end{equation}
where $P_{\rm NT}$ is non-thermal pressure and $P$ is thermal pressure.

@@

% Starburst 
In the starburst phase, MCCs have velocity dispersion, sizes,
and virial parameter comparable to those in the accreting phase; however, the MCCs in the starburst phase span 
a wider range in $\sigma^2/R$ and gas surface density. 
Due to a time delay between cloud collapse and \SF, we also examine the virial parameter of MCCs in the pre-starburst phase. 
As shown in \Fig{alpha21-22}, the total virial parameter (blue dots in the figure) 
of a subset of MCCs in the pre-starburst phase are lower than those in the starburst phase of the same mass range, which is as expected.
%Similarly, the stellar-to-gas mass ratio of MCCs in the starburst phase is higher overall (see bottom panels of \Fig{alpha21-22}).
%These trends are as expected in the MCC properties prior to and during a starburst. 
% None of these MCCs are in the unstable regime ($alpha_{\rm vir} < 2$) because the MCCs are not directly the collapsing structures, 
% i.e., we haven’t reached those scales where true collapse is happening. 
% This implication of this is that to examine the truly star-forming structures (cores) will require resolution smaller than ~40 pc.
As \SF proceed, stellar feedback inject kinetic energy in the form of turbulence. As shown in 
\Fig{Mach_stellarRatio}, the Mach number of MCCs with a stellar-to-gas mass ratio of $M_\star/M_{\rm gas} \simeq 0.1 - 1$ 
(i.e., excluding the main component of \flower with $M_\star/M_{\rm gas} \gtrsim 1$; see e.g., \Fig{stellarRatio16})
increases from the pre-starburst phase through the post-starburst phase.

@@

Excluding the biggest MCC identified in each evolutionary stage (i.e., the top right cluster of points in the 
linewidth-size relation of \Fig{larsons_single} and the top left panel of \Fig{alpha16-28}), which correspond to 
the main disk of \flower, the MCCs of \flower have dynamics largely similar to those observed in
$z$\ssim2 spatially resolved studies of gas-rich star-forming
galaxies, in terms of their velocity dispersions, sizes, and gas
surface densities (\Fig{larsons_single}; see e.g.,
\citealt{Swinbank11a}), with sizes of the order of $R\simeq$\,100\,pc
and velocity dispersions of $\sigma\simeq$\,20$-$80\,\kms. 
Their velocity dispersions are also comparable to those observed in the inner Milky way and nearby gas-rich galaxies (e.g., M64; \citealt{Oka01a, Rosolowsky05a, Heyer09a}), which lies along the locus of $\sigma\propto R^{0.56}$. Such high velocity dispersions and surface densities are expected since they are located in the nuclear regions of the galaxy, where the potential well is also deep. The higher velocity dispersions
(see footnote~\ref{ftn:veldisp} on page \pageref{ftn:veldisp}) of
these MCCs can also be understood as they have experienced more recent
episodes of \SF as \flower is assembling its stellar mass --- these MCCs have higher stellar-to-gas mass ratios of $M_\star/M_{\rm gas}\simeq0.1 - 1$ (\Fig{stellarRatio16}) and high turbulent pressure as shown by the high $\mathcal{M}$ in \Fig{Mach_stellarRatio}.  
%
In fact, by the accreting stage (see \Fig{SFH}a) at $z\simeq$\,7.2, 
\flower has assembled a stellar mass of
$M_\star$\eq7.5\E{9}\,\Msun. Thus, the higher velocity dispersion is
due in part to the stronger stellar feedback \MM{Again, not clear that
  this is feedback and not potential driven. Can you demonstrate one
  or the other explicitly?}

The high pressure observed in \flower \MM{Where are pressures shown?
  You can't just draw them from the virial vs.\ surface density plot
  because that assumes equilibrium gas clouds, which these are clearly
  not, given their huge stellar contents.} is comparable to what has been observed in local ultra-luminous IR galaxies (ULIRGs); however, the molecular clouds in the latter are concentrated within their central regions and have typical sizes of $\sim$70-100\,pc and masses on the order of $\sim$10$^9$\,\Msun \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCCs of \flower are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
For instance, in local ULIRGs, 
%mm they are likely form 
  MCCs are likely formed
by shock compression and cloud-cloud collisions 
%mm after large amounts of gas from the pair of progenitor galaxies
%are funneled toward the central region 
    that funnel large amounts of gas from the progenitor galaxies
    towards the central region
\citep{Tan00a, Wu18a}.
% Tasker09a
On the other hand, the highly turbulent structures of \flower likely result from 
extra-planar flows that are due to gas accretion onto the galaxy and SN-driven outflows \citep{gallerani:2018,kohandel:2019}.
The presence of such extra-planar flows may \MM{Do we have any
  evidence for or against this speculation?} also be the 
dominant mode for forming the highly supersonic massive MCCs observed
in gas-rich star-forming galaxies at $z$\ssim2 (see also e.g.,
\citealt{Swinbank11a}) --- but the observed size scales could also be biased 
high due to the resolution limit.
The constant inter-cloud collisions as gas is being accreted onto the
main galaxy 
%\MM{This is a highly dissipative process.  Hennebelle \&
%  Klessen found it could only be important in rather small galaxies.
%  Does \flower qualify?} 
would also render the MCCs gravitationally unbound (see e.g., \citealt{Dobbs11a}), 
%\MM{Dobbs
%  started with a low-density disk, and got low-density clouds. I don't
%  think that's what you have here.  Do you have any evidence for
%  inter-cloud collisions being important?}
%\DL{think okay to state as speculation since we are not drawing conclusions based on this?} 
which would be consistent with the high $\alpha_{\rm vir}$ found for the MCCs in the main disk of \flower.

@@

The largest molecular structure identified is essentially the main
disk of \flower \MM{Is this consistent with your assertion of no shear
  in a footnote above?} and breaks down into smaller substructures that are
denser as we increase $n_{\rm cut}$ (see e.g.,
\Fig{alpha16-28}). Overall, the MCCs we identified are much bigger in
size and mass than nearby giant molecular clouds (GMCs), but comparable to those observed in $z$\,\ssim2$-$4 galaxies based on spatially resolved imaging \citep{Swinbank11a}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.9\textwidth]{\figpath/{ss16_toomre_combined_2by2_0_2.0}.pdf}
\caption{
Surface density maps of the gas (top left) and stellar (top right) components of \flower and
their radial velocity dispersion maps projected onto the $xy$-plane.
%
\label{fig:sigma}}
\end{figure*}

@@

Note that even in stages when \flower~is star forming we find that most of its MCCs have high $\alpha_{\rm vir}$. This seem to be self-contradictory: if most of the molecular gas in \flower are stable against collapse, how does it sustain its high SFR of $\sim$70\,\Msun\,yr\pmOne in the starburst phase? Such tension in the interpretation is understood as follows.
%
First, our results are limited by the resolution of the simulation, i.e., in \flower~we do not resolve \SF in {\it cores} and {\it clumps} as in the classical molecular cloud hierarchy. This re-iterates the statement about low $Q$ above.
%
Second, due to turbulence dissipation on small-scales, such
(sub-)structures are no longer supported by large-scale gravitational
potential and differential rotation.  \MM{Not sure how potential can
  support, and in that footnote above you asserted that shear was
  unimportant, contradicting this claim.} This in turn enables sub-regions of the MCCs to collapse and potentially form OB associations out of these unbound MCCs \citep{Clark04a, Clark05a}.
%
Further, mechanisms causing non-axisymmetric perturbations (e.g.,
arms) to the gravitational potential \MM{Gravitational instability is
  that mechanism (see e.g.\ Wada et al.'s classic review of the
  subject (2011, doi:10.1017/S1743921311000640), so this isn't really a separate mechanism.} can induce orbit crossing, shocks and dissipation in gas, promoting turbulence dissipation and collapse on smaller scales to trigger \SF. These mechanisms are likely important given high $Q_{\rm eff}$ found across \flower and its few tens of solar masses per year SFR, albeit sporadic.
Finally, the SFR of \flower is very bursty and high SFR stages are only transient.
% the giant MCCs found are likely due to non-axisymmetric mode.

@@

\MM{Each item in the summary needs to refer to definite results and
  analysis in the previous paper.  I have noted the items that don't
  seem to me to have that necessary support.}
We study the dynamical properties of molecular clouds complexes and their temporal evolution in a \z$\sim$\,6 prototypical galaxy, \flower,
at the EoR using the state-of-the-art cosmological zoom-in simulation (\ncode{Serra}),
which includes a chemical network to determine the formation of molecular
hydrogen, heating and cooling of the ISM by
%mm metals
    UV radiation and metal lines, 
and detailed stellar feedback.
We use a clump-finding algorithm and a set of H$_2$ volumetric densities
to identify MCCs and their sub-structures in \flower and its satellites.
We decompose the molecular structures into non-overlapping 
%mm tiles
     objects
by identifying a set of 
%mm different density contours at different 
     density contours with varying cutoff densities at multiple
evolutionary stages.
Such an identification method is 
%mm similar 
     the three-dimensional analog
to MCC identification based on 
%mm contour levels/surfaces of surface brightness maps/cubes 
    intensity levels in surface brightness maps or
    position-position-velocity cubes
of molecular line tracers (e.g., CO, CS, NH$_3$),
since the line luminosity scales with the molecular gas density (modulo optical depth effects).

We extract properties such as mass, size, Mach number, velocity
dispersion, gas surface density, and virial parameter for each MCC and
we perform a Toomre-$Q$ stability analysis
%mm
    on the galaxy. 
The properties of the MCCs in \flower are compared with observed MCCs in the Milky Way disk, the Galactic center, and gas-rich starburst galaxies in the local Universe and at the peak epoch of cosmic \SF. This is done for different evolutionary stages of \flower.

We find that the MCCs of \flower 
%mm are highly supersonic, with high 
   have highly supersonic
velocity dispersions ($\sigma\simeq$\,200\,\kms) comparable to
those observed in $z$\ssim2 starburst galaxies.
The mass scale of the MCCs is of the order of $10^{5.5-8.5}$\,\Msun,
depending on the chosen H$_{2}$ density cut. 
The $\sim$200\,pc-scale MCCs found with a low density threshold
correspond to the arms of the disk of \flower. These break down into smaller $\lesssim$100\,pc-scale MCCs at higher density thresholds.
The more massive and bigger MCCs in \flower compared to the Milky Way
likely result from the higher gas mass fraction, surface density, and velocity dispersion,
which set the scale for fragmentation. \MM{This could be quantified in
  the discussion by calculating the Jeans scale for the disk conditions.}
This physical scale is
consistent with what has been found in higher resolution idealized
simulation of isolated galaxies. \MM{Perhaps this could be shown more explicitly?}
Our simulation here confirms these results at EoR and under
the influence of continuous gas accretion from the surrounding IGM (see e.g., \citealt{Klessen10a, Goldbaum11a}).

>>>>

The local sound speed of all identified molecular complexes is much
smaller than their turbulent velocities\footnote{Velocity dispersions
  calculated from velocity field are comparable to those calculated
  from the pressure field 
    \MM{The turbulent pressure field has not been
    mentioned in section 3, so you'll need to add a sentence there
    mentioning its presence and behavior to mention it here}, 
    \DL{AP: also mention turbulent pressure found for the bulk of dense gas in the simulation in the phase-space analysis presented in \citet{Pallottini17a} here?}
indicating that contribution from shear is unlikely to dominate the velocity dispersions of the smaller MCCs identified, especially 
at the highest $n_{\rm ncut}$.\label{ftn:veldisp}}. 
On average, the Mach number of the MCCs identified at the highest density threshold is $\mathcal{M} = \sigma_{\rm NT}/c_s\simeq 50$, 
but this depends on the evolutionary stage of \flower.  % see e.g., test_brute/leaf_fields_ss16-28/18.96_10_fields/Mach_cloud-mass.png
This is consistent with the global analysis done on \flower by \citet{Vallini18a}.
We also compute the Mach number from turbulent and thermal pressure to determine the importance of 
differential rotation in the derived velocity dispersion: 
\begin{equation}
\mathcal{M} \equiv \sqrt{\frac{P_{\rm NT} + P}{ P }}\,,
\label{eqn:MPress}
\end{equation}
where $P_{\rm NT}$ is non-thermal pressure and $P$ is thermal pressure.

@@

% Starburst 
In the starburst phase, MCCs have velocity dispersion, sizes,
and virial parameter comparable to those in the accreting phase; however, the MCCs in the starburst phase span 
a wider range in $\sigma^2/R$ and gas surface density. 
Due to a time delay between cloud collapse and \SF, we also examine the virial parameter of MCCs in the pre-starburst phase. 
As shown in \Fig{alpha21-22}, the total virial parameter (blue dots in the figure) 
of a subset of MCCs in the pre-starburst phase are lower than those in the starburst phase of the same mass range, which is as expected.
%Similarly, the stellar-to-gas mass ratio of MCCs in the starburst phase is higher overall (see bottom panels of \Fig{alpha21-22}).
%These trends are as expected in the MCC properties prior to and during a starburst. 
% None of these MCCs are in the unstable regime ($alpha_{\rm vir} < 2$) because the MCCs are not directly the collapsing structures, 
% i.e., we haven’t reached those scales where true collapse is happening. 
% This implication of this is that to examine the truly star-forming structures (cores) will require resolution smaller than ~40 pc.
As shown in \Fig{Mach_stellarRatio}, the Mach number (see Equation~\ref{eqn:MPress}) 
of MCCs with a stellar-to-gas mass ratio of $M_\star/M_{\rm gas} \simeq 0.1 - 1$ 
(i.e., excluding the main component of \flower with $M_\star/M_{\rm gas} \gtrsim 1$; see e.g., \Fig{stellarRatio16})
increases from the pre-starburst phase through the post-starburst phase. 
This results as stellar feedback injects kinetic energy into the surrounding ISM in the form of turbulence as \SF proceed.
In the quiescent phase, high Mach number are seen in mostly MCCs identified with the lowest density threshold. These 
correspond to satellites MCCs in the outskirts of \flower.

@@

Excluding the biggest MCC identified in each evolutionary stage (i.e., the top right cluster of points in the 
linewidth-size relation of \Fig{larsons_single} and the top left panel of \Fig{alpha16-28}), which correspond to 
the main disk of \flower, the MCCs of \flower have dynamics largely similar to those observed in
$z$\ssim2 spatially resolved studies of gas-rich star-forming
galaxies, in terms of their velocity dispersions, sizes, and gas
surface densities (\Fig{larsons_single}; see e.g.,
\citealt{Swinbank11a}), with sizes of the order of $R\simeq$\,100\,pc
and velocity dispersions of $\sigma\simeq$\,20$-$80\,\kms. 
Their velocity dispersions are also comparable to those observed in the inner Milky way and nearby gas-rich galaxies (e.g., M64; \citealt{Oka01a, Rosolowsky05a, Heyer09a}), which lies along the locus of $\sigma\propto R^{0.56}$. Such high velocity dispersions and surface densities are expected since they are located in the nuclear regions of the galaxy, where the potential well is also deep. The higher velocity dispersions
(see footnote~\ref{ftn:veldisp} on page \pageref{ftn:veldisp}) of
these MCCs can also be understood as they have experienced more recent
episodes of \SF as \flower is assembling its stellar mass --- these MCCs have higher stellar-to-gas mass ratios of $M_\star/M_{\rm gas}\simeq0.1 - 1$ (\Fig{stellarRatio16}) and high turbulent pressure as shown by the high $\mathcal{M}$ in \Fig{Mach_stellarRatio}.  
%
In fact, by the accreting stage (see \Fig{SFH}a) at $z\simeq$\,7.2, 
\flower has assembled a stellar mass of
$M_\star$\eq7.5\E{9}\,\Msun.

The high pressure observed in \flower (see \Fig{Mach_stellarRatio}, see also \citealt{Pallottini17a})
is comparable to what has been observed in local ultra-luminous IR galaxies (ULIRGs); however, the molecular clouds in the latter are concentrated within their central regions and have typical sizes of $\sim$70-100\,pc and masses on the order of $\sim$10$^9$\,\Msun \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCCs of \flower are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
For instance, in local ULIRGs, 
%mm they are likely form 
  MCCs are likely formed
by shock compression and cloud-cloud collisions 
%mm after large amounts of gas from the pair of progenitor galaxies
%are funneled toward the central region 
    that funnel large amounts of gas from the progenitor galaxies
    towards the central region
\citep{Tan00a, Wu18a}.
% Tasker09a
On the other hand, the highly turbulent structures of \flower likely result from 
extra-planar flows that are due to gas accretion onto the galaxy and SN-driven outflows \citep{gallerani:2018,kohandel:2019}.
The presence of such extra-planar flows may \MM{Do we have any
  evidence for or against this speculation?} also be the 
dominant mode for forming the highly supersonic massive MCCs observed
in gas-rich star-forming galaxies at $z$\ssim2 (see also e.g.,
\citealt{Swinbank11a}) --- but the observed size scales could also be biased 
high due to the resolution limit.
The constant inter-cloud collisions as gas is being accreted onto the
main galaxy 
%\MM{This is a highly dissipative process.  Hennebelle \&
%  Klessen found it could only be important in rather small galaxies.
%  Does \flower qualify?} 
would also render the MCCs gravitationally unbound (see e.g., \citealt{Dobbs11a}), 
%\MM{Dobbs
%  started with a low-density disk, and got low-density clouds. I don't
%  think that's what you have here.  Do you have any evidence for
%  inter-cloud collisions being important?}
%\DL{think okay to state as speculation since we are not drawing conclusions based on this?} 
which would be consistent with the high $\alpha_{\rm vir}$ found for the MCCs in the main disk of \flower.

@@

The largest molecular structure identified is essentially the main
disk of \flower \MM{Is this consistent with your assertion of no shear
  in a footnote above?} and breaks down into smaller substructures that are
denser as we increase $n_{\rm cut}$ (see \Sec{ncut}). Overall, the MCCs we identified are much bigger in
size and mass than nearby giant molecular clouds (GMCs), but comparable to those observed in $z$\,\ssim2$-$4 galaxies based on spatially resolved imaging \citep{Swinbank11a}.

@@

\begin{figure*}
\centering
\includegraphics[width=0.9\textwidth]{\figpath/{ss16_toomre_combined_2by2_0_2.0}.pdf}
\caption{
Surface density maps of the gas (top left) and stellar (top right) components of \flower and
their radial velocity dispersion maps projected onto the $xy$-plane.
Center of mass positions of MCCs within $\sim$1.5\,kpc of \flower identified with $n_{\rm cut}$\eq6.81\,\cc are overplotted as star symbols. 
\label{fig:sigma}}
\end{figure*}

@@

Note that even in stages when \flower~is star forming we find that most of its MCCs have high $\alpha_{\rm vir}$. This seem to be self-contradictory: if most of the molecular gas in \flower are stable against collapse, how does it sustain its high SFR of $\sim$70\,\Msun\,yr\pmOne in the starburst phase? Such tension in the interpretation is understood as follows.
%
First, our results are limited by the resolution of the simulation, i.e., in \flower~we do not resolve \SF in {\it cores} and {\it clumps} as in the classical molecular cloud hierarchy. This re-iterates the statement about low $Q$ above.
%
Second, due to turbulence dissipation on small-scales, such
(sub-)structures are no longer supported by differential rotation.  \MM{footnote above, you asserted that shear was
  unimportant, contradicting this claim.} This in turn enables sub-regions of the MCCs to collapse and potentially form OB associations out of these unbound MCCs \citep{Clark04a, Clark05a}.
%
Further, mechanisms causing non-axisymmetric perturbations (e.g.,
arms) to the gravitational potential \MM{Gravitational instability is
  that mechanism (see e.g.\ Wada et al.'s classic review of the
  subject (2011, doi:10.1017/S1743921311000640), so this isn't really a separate mechanism.} can induce orbit crossing, shocks and dissipation in gas, promoting turbulence dissipation and collapse on smaller scales to trigger \SF. These mechanisms are likely important given high $Q_{\rm eff}$ found across \flower and its few tens of solar masses per year SFR, albeit sporadic.
Finally, the SFR of \flower is very bursty and high SFR stages are only transient.
% the giant MCCs found are likely due to non-axisymmetric mode.

@@

\MM{Each item in the summary needs to refer to definite results and
  analysis in the previous paper.  I have noted the items that don't
  seem to me to have that necessary support.}
We study the dynamical properties of molecular clouds complexes and their temporal evolution in a \z$\sim$\,6 prototypical galaxy, \flower,
at the EoR using the state-of-the-art cosmological zoom-in simulation (\ncode{Serra}),
which includes a chemical network to determine the formation of molecular
hydrogen, heating and cooling of the ISM by
    UV radiation and metal lines, 
and detailed stellar feedback.
We use a clump-finding algorithm and a set of H$_2$ volumetric densities
to identify MCCs and their sub-structures in \flower and its satellites.
We decompose the molecular structures into non-overlapping 
     objects
by identifying a set of 
     density contours with varying cutoff densities at multiple
evolutionary stages.
Such an identification method is 
     the three-dimensional analog
to MCC identification based on 
    intensity levels in surface brightness maps or
    position-position-velocity cubes
of molecular line tracers (e.g., CO, CS, NH$_3$),
since the line luminosity scales with the molecular gas density (modulo optical depth effects).

We extract properties such as mass, size, Mach number, velocity
dispersion, gas surface density, and virial parameter for each MCC and
we perform a Toomre-$Q$ stability analysis on the galaxy. 
The properties of the MCCs in \flower are compared with observed MCCs in the Milky Way disk, the Galactic center, and gas-rich starburst galaxies in the local Universe and at the peak epoch of cosmic \SF. This is done for different evolutionary stages of \flower.

We find that the MCCs of \flower 
have highly supersonic
velocity dispersions ($\sigma\simeq$\,200\,\kms) comparable to
those observed in $z$\ssim2 starburst galaxies.
The mass scale of the MCCs is of the order of $10^{5.5-8.5}$\,\Msun,
depending on the chosen H$_{2}$ density cut. 
The $\sim$200\,pc-scale MCCs found with a low density threshold
correspond to the arms of the disk of \flower. These break down into smaller $\lesssim$100\,pc-scale MCCs at higher density thresholds.
The more massive and bigger MCCs in \flower compared to the Milky Way
likely result from the higher gas mass fraction, surface density, and velocity dispersion,
which set the scale for fragmentation. \MM{This could be quantified in
  the discussion by calculating the Jeans scale for the disk conditions.}
This physical scale is
consistent with what has been found in higher resolution idealized
simulation of isolated galaxies. \MM{Perhaps this could be shown more explicitly?}
Our simulation here confirms these results at EoR and under
the influence of continuous gas accretion from the surrounding IGM (see e.g., \citealt{Klessen10a, Goldbaum11a}).
